---
title: 'GEOG606: Class Project'
author: "Anika Cartas"
date: "Sunday, April 26, 2015"
output: html_document
---

1. Determine how many samples are needed to control the estimation error within 20%, 10%, and 5% of the total biomass

```{r}
#install.packages("raster")
library(raster)

truth <- raster("class_project/Class_Project_NC_biomass")

N <- ncol(truth) * nrow(truth)
sd <- sd(truth[1:N])
mean <- mean(truth[1:N])
CV <- sd/mean #slide 7, bullet point 2, Prof. Huang said this is it.

#This formula is from end of slide 7, lecture 11. 
get_sample_size <- function(AE, N, CV){ 
  n <-  (4*N*CV^2) / (N*AE^2 + 4*CV^2) 
}

n.20 <- get_sample_size(0.2,N,CV)
n.10 <- get_sample_size(0.1,N,CV)
n.5  <- get_sample_size(0.05,N,CV)
```

2. Select pixels from the reference biomass map and use them to estimate the total biomass and its standard error

Once we know how many samples to take, we can use raster library to randomly sample from our files. 
```{r}
height <- stack("class_project/Class_Project_NC_lidar_profiles10") #stores all 10 bands

#FYI, you can access each rasterLayer in the rasterStack individually.
#h.10 <- height@layers[[1]] #first of 10 bands, etc.
#plot(h.10, main="Lidar Profiles 1")

get_sample_data <- function(n){
  samp <- sample(N,n)     #Cell number of our samples
  bio <- truth[samp]      #Values at each sample cell
  hgt <- height[samp]
  data <- as.data.frame(cbind(bio,hgt))
  colnames(data) <- c("Biomass","H1","H2","H3","H4","H5","H6","H7","H8","H9","H10")

  data #return the df
}

data.5 <- get_sample_data(n.5)
data.10 <- get_sample_data(n.10)
data.20 <- get_sample_data(n.20)

```

Stats Group: Stepwise linear regression

Model group 1: Use the samples selected to derive sample based estimates with a relative error of 5% and 20% to build your model

Below is a demo assuming we found the right number for 20% error (n.20)
```{r}
attach(data.20) #Not sure how to do this without attach/detach
lm.20 <- step(lm(Biomass ~ 1), as.formula(Biomass ~ H1+H2+H3+H4+H5+H6+H7+H8+H9+H10), direction="forward")
lm.20
detach(data.20)

attach(data.10) #Not sure how to do this without attach/detach
lm.10 <- step(lm(Biomass ~ 1), as.formula(Biomass ~ H1+H2+H3+H4+H5+H6+H7+H8+H9+H10), direction="forward")
lm.10
detach(data.10)

attach(data.5) #Not sure how to do this without attach/detach
lm.5 <- step(lm(Biomass ~ 1), as.formula(Biomass ~ H1+H2+H3+H4+H5+H6+H7+H8+H9+H10), direction="forward")
lm.5
detach(data.5)

```

Model group 2: Randomly draw 2 curves across the study area as two faked roads. Use samples along the roads to build your model

```{r}

sample_road <-function(n,plot=F){
  if(plot==T){
    plot(truth,main="Biomass (Mg/ha)")
    curve(5000*sin(0.0001*x)+620000,from=848171,to=887541,add=T,n=2000) #Road 1
    curve(3000*cos(0.0001*x)+625000,from=848171,to=887541,add=T,n=2000) #Road 2
  }
  
  #For AE = 20%
  samp.road.x <- sample(c(848171:887541),n) #get x values to sample at.
  samp.road.y <- sapply(samp.road.x, function(x){
    if (rbinom(1,1,0.5) == 0) #Coin flip which road we use
      y <- 5000*sin(0.0001*x)+620000
    else
      y <- 3000*cos(0.0001*x)+625000
    }) #matching y values
  
  if (plot==T){
    points(x=samp.road.x,y=samp.road.y,col="blue")
  }

  samp.road <- cbind(x=samp.road.x,y=samp.road.y)

  #Use extract function to get values at given x/y
  bio.road <- extract(truth,samp.road)
  height.road <- extract(height,samp.road)
  data.road <- as.data.frame(cbind(bio.road,height.road))
  colnames(data.road) <- c("Biomass","H1","H2","H3","H4","H5","H6","H7","H8","H9","H10")

  data.road
}

data.road.5 <- sample_road(n.5,plot=T)
data.road.20 <- sample_road(n.20,plot=T)

```

Calculate total biomass
Apply the developed models to the entire study area to produce biomass maps
Calculate the total biomass from each map

Comparing the biomass estimates derived using different methods against the “truth” derived from the reference map.

```{r}

(biomass.truth <- sum(truth[1:80000])) #Total biomass of truth.
#biomass.srs.20 <- 
#biomas.srs.5 <- 
#biomass.road.20 <- 
#biomass.road.5 <- 
```
