---
title: 'GEOG606: Final Presentation'
author: "Anika Cartas"
date: "Thursday, May 07, 2015"
output: html_document
---

```{r read_data}
library(ggplot2)
data <- read.csv("results_v4.csv")
#write.csv(data,"results_v4.csv")
```

**Calculate FRBWidth.State**

Two windows in a row need to be non-forest to be cut-off.

```{r, echo=FALSE, eval=FALSE}
TIDS <- unique(data$TransectID)
for (id in TIDS){
  sub <- data[data$TransectID == id,]
  FRBWidth <- 0
  done <- FALSE
  
  while (!done & FRBWidth < 200){
    if (sub[sub$Distance == FRBWidth + 10,"LU"] == 40){
      FRBWidth <- FRBWidth + 10
    } 
    else 
      if (FRBWidth < 190 && sub[sub$Distance == FRBWidth + 20,"LU"] == 40){
        FRBWidth <- FRBWidth + 20
      } 
      else {
        done <- TRUE
      }
  }
  
  data[data$TransectID == id,"FRBWidth.State"] <- FRBWidth
  cat("TID",id,"",FRBWidth,"\n")

}
write.csv(data,"results_v4.csv")

```

**Recalculate FRBWidth.Lidar**

Two windows in a row (20m) need to be <30% in order to cut off the FRB.

```{r, echo=FALSE,eval=FALSE}
#Don't run again.
TIDS <- unique(data$TransectID)
for (id in TIDS){
  sub <- data[data$TransectID == id,]
  addDist <- 10
  FRBWidth <- sub$FRBWidth.Lidar[1]
  
  done <- FALSE
  
  while (!done && FRBWidth < 190){
    cur <- sub[sub$Distance == FRBWidth + 10,]
    nex <- sub[sub$Distance == FRBWidth + 20,]
    
    if (nex$Cover >= 0.3){
      FRBWidth <- FRBWidth + 20
    }
    else {
      done <- TRUE
    }
  }
  
  data[data$TransectID == id,"FRBWidth.Lidar"] <- FRBWidth
  if (FRBWidth != sub$FRBWidth.Lidar[1]){
    cat("TID",id,"Old",sub$FRBWidth.Lidar[1],"New",FRBWidth,"\n")
  }
}
write.csv(data,"results_v4.csv")

```

**1. Summary Statistics on FRBWidth Lidar/State**

```{r}
t.first <- data[data$Id == 1, ]
FRBWidth.Lidar.mean <- mean(t.first$FRBWidth.Lidar)
FRBWidth.State.mean <- mean(t.first$FRBWidth.State)

cat("===FRBWidth===\nMean:\n\tLidar:",FRBWidth.Lidar.mean,"\n\tState:",FRBWidth.State.mean,
    "\nSD:\n\tLidar:",sd(t.first$FRBWidth.Lidar),"\n\tState:",sd(t.first$FRBWidth.State),
    "\nVar:\n\tLidar:",var(t.first$FRBWidth.Lidar),"\n\tState:",var(t.first$FRBWidth.State),"\n")

hist(t.first$FRBWidth.Lidar, probability=TRUE, breaks=seq(-10,200,10), main="Histogram of FRBWidth",xlab="FRBWidth",ylim=c(0,0.035),col=rgb(1,0,0,0.25))
hist(t.first$FRBWidth.State, probability=TRUE, breaks=seq(-10,200,10), add=T, col=rgb(0,0,1,0.25))

legend("top",c("Lidar","State"),col=c(rgb(1,0,0,0.5),rgb(0,0,1,0.5)),lwd=c(2.5,2.5))
```

**2a. FRBWidth based on immediately following land-use.**

Probably not the best method. Should check more directly on ratio of following stuff.

```{r}
#Followed by Agriculture: Distance = FRBWidth + 10 AND LU = 20
t.ag <- data[data$Distance == (data$FRBWidth.State+10) & data$LU == 20, ]
t.ag.mean <- mean(t.ag$FRBWidth.State)
t.ag.sd <- sd(t.ag$FRBWidth.State)
t.ag.var <- var(t.ag$FRBWidth.State)

#Followed by Developed
t.dev <- data[data$Distance == (data$FRBWidth.State+10) & data$LU == 10, ]
t.dev.mean <- mean(t.dev$FRBWidth.State)
t.dev.sd <- sd(t.dev$FRBWidth.State)
t.dev.var <- var(t.dev$FRBWidth.State)

#Followed by Forest
t.for <- data[data$Distance == (data$FRBWidth.State+10) & data$LU == 40, ]
t.for.mean <- mean(t.for$FRBWidth.State)
t.for.sd <- sd(t.for$FRBWidth.State)
t.for.var <- var(t.for$FRBWidth.State)
```
```{r, echo=FALSE}
cat("Immediately by Ag:\n\tMean =",t.ag.mean,"\n\tSD =",t.ag.sd,"\n\tVar =",t.ag.var)
cat("Immediately by Dev:\n\tMean =",t.dev.mean,"\n\tSD =",t.dev.sd,"\n\tVar =",t.dev.var)
cat("Immediately by Forest:\n\tMean =",t.for.mean,"\n\tSD =",t.for.sd,"\n\tVar =",t.for.var)
```

**Comparing Histograms**

```{r, echo=FALSE}
par(mfrow=c(2,2))
hist(t.first$FRBWidth.Lidar,  col=rgb(0.8,0.8,0.8),probability=TRUE, breaks=20, main="All Transects",xlab="FRBWidth",ylim=c(0,0.02),xlim=c(0,200))
hist(t.ag$FRBWidth.Lidar, col=rgb(1,1,.7), probability=TRUE,breaks=20,main="Immediately by Ag",xlab="FRBWidth",ylim=c(0,0.02),xlim=c(0,200))
hist(t.dev$FRBWidth.Lidar, col=rgb(1,192/255,203/255), probability=TRUE,breaks=20,main="Immediately by Dev",xlab="FRBWidth",ylim=c(0,0.02),xlim=c(0,200))
hist(t.for$FRBWidth.Lidar, col=rgb(60/255,179/255,113/255), probability=TRUE,breaks=20,main="Immediately by Forest",xlab="FRBWidth",ylim=c(0,0.02),xlim=c(0,200))
```

**2b. Comparing FRB of most frequent following LU**


```{r,eval=FALSE, echo=FALSE}
#Don't run this again, I had to make 3 manual adjustments for 104 to pick the best fit for nextLU

#For each transect id
  #Check LU of all with Dist > FRBWidth
  #Find median value
  #Assign as Frequent-Follower for whole id.

data$Next.LU.Alt <- NA

Mode <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux)); ux[tab == max(tab)]
}

getNextLU <- function(id){
  #Test on subset
  sub <- data[data$TransectID == id,]
  past <- sub[sub$Distance > sub$FRBWidth,]
  if (nrow(past) > 0){
    nextLU <- Mode(past$LU)  
  }
  else{
    nextLU <- NA
  }
  
  nextLU #return value

}

for (id in unique(data$TransectID)){
  nextLU <- getNextLU(id)
  if (length(nextLU) > 1){
    if (nextLU[1] == 40){
      data[data$TransectID == id,"Next.LU"] <- nextLU[2]
      data[data$TransectID == id,"Next.LU.Alt"] <- nextLU[1]    
    }
    else if (nextLU[2] == 40){
      data[data$TransectID == id,"Next.LU"] <- nextLU[1]
      data[data$TransectID == id,"Next.LU.Alt"] <- nextLU[2]          
    }
    else {
      cat("TID",id,"choose a mode",nextLU,"\n")      
    }
  }
  else{
    data[data$TransectID == id,"Next.LU"] <- nextLU
  }
  
}

write.csv(data,"results_v4.csv")
```

```{r,echo=FALSE}
#Followed by Agriculture: Distance = FRBWidth + 10 AND LU = 20
t.ag <- data[data$Distance == (data$FRBWidth+10) & data$Next.LU == 20 & !is.na(data$Next.LU), ]
t.ag.mean <- mean(t.ag$FRBWidth)
t.ag.sd <- sd(t.ag$FRBWidth)
t.ag.var <- var(t.ag$FRBWidth)

#Followed by Developed
t.dev <- data[data$Distance == (data$FRBWidth+10) & data$Next.LU == 10 & !is.na(data$Next.LU), ]
t.dev.mean <- mean(t.dev$FRBWidth)
t.dev.sd <- sd(t.dev$FRBWidth)
t.dev.var <- var(t.dev$FRBWidth)

#Followed by Forest
t.for <- data[data$Distance == (data$FRBWidth+10) & data$Next.LU == 40 & !is.na(data$Next.LU), ]
t.for.mean <- mean(t.for$FRBWidth)
t.for.sd <- sd(t.for$FRBWidth)
t.for.var <- var(t.for$FRBWidth)
```

```{r, echo=FALSE}
cat("Followed by Ag:\n\tMean =",t.ag.mean,"\n\tSD =",t.ag.sd,"\n\tVar =",t.ag.var)
cat("Followed by Dev:\n\tMean =",t.dev.mean,"\n\tSD =",t.dev.sd,"\n\tVar =",t.dev.var)
cat("Followed by Forest:\n\tMean =",t.for.mean,"\n\tSD =",t.for.sd,"\n\tVar =",t.for.var)
```

**Comparing Histograms**

```{r, echo=FALSE}
par(mfrow=c(2,2))
hist(t.first$FRBWidth,  col=rgb(0.8,0.8,0.8),probability=TRUE, breaks=20, main="All Transects",xlab="FRBWidth",ylim=c(0,0.05),xlim=c(0,200))
hist(t.ag$FRBWidth, col=rgb(1,1,.7), probability=TRUE,breaks=20,main="Followed by Ag",xlab="FRBWidth",ylim=c(0,0.05),xlim=c(0,200))
hist(t.dev$FRBWidth, col=rgb(1,192/255,203/255), probability=TRUE,breaks=20,main="Followed by Dev",xlab="FRBWidth",ylim=c(0,0.05),xlim=c(0,200))
hist(t.for$FRBWidth, col=rgb(60/255,179/255,113/255), probability=TRUE,breaks=20,main="Followed by Forest",xlab="FRBWidth",ylim=c(0,0.05),xlim=c(0,200))
```

**3. Height Statistics based on FRBWidth**

None of this stuff really tells me anything yet, not sure what I'm trying to answer with it.

```{r}
heightStats <- data.frame(Mean.Height=numeric(200),Distance=numeric(200),FRBWidth=numeric(200))

i <- 1
width <- 0

while (i <= 200){
  width <- width+1
  n <- 1
  while (n <= width){
    dist <- n*10
    heightStats[i,] <- cbind(mean(data[data$FRBWidth == width*10 & data$Distance == dist,"Height"]),dist,width*10)
    i<-i+1
    n<-n+1
  }
}

plot(x=heightStats$Distance,y=heightStats$Mean.Height,xlab="Distance",ylab="Mean Height",col=as.factor(heightStats$FRBWidth), pch="+")

plotAtDistance <- function(d){
  sub <- heightStats[heightStats$Distance == d,]
  label <- paste("Distance =",d)
    plot(y=sub$Mean.Height, x=sub$FRBWidth,ylab="Mean Height",xlab="FRB Width",main=label,pch=20,type="o",ylim=c(0,20),xlim=c(0,200))
  
  sub #return df
}
```

**Printing out a bunch of plots**

```{r, echo=FALSE}
par(mfrow=c(1,2))
for (i in seq(10,200,10)){
  plotAtDistance(i)
}
```