---
title: "GEOG606 - HW1"
author: "Anika Cartas"
date: "Saturday, February 14, 2015"
output: word_document
---

#1. Solve the following linear equations:

-x + 2y = 1
3x +  y = 3

```{r}

```

If we add another equation, x + y = 2, is there a solution for the set of 3 equations? If not, could you give a best estimate of x and y that satisfy the equations?

```{r}
```

#2. Calculate the covariance of the following 3*2 data matrix. 

```{r}
```

What does the covariance tell you about the data?

#3.Use the folowing dataset to demonstrate the Central Tendency Theorem.

3.1. Randomly select 5 values and calculate their mean. Repeat 3 times, and calculate the mean of the sampling mean. Then, repeat 5, 10, 20, adn 50 times. Use a plot to show how the mean of the sampling mean changes as the number of repeats increases.

3.2. Repeat 3.1., but select 10 values during each random selection. Explain these plots in relation to the Central Tendency Theorem.

3.3. How does the sampling mean approach normal distribution in these experiments? Use histograms to show.

```{r}
data = c(3,2,3,4,3,2,5,4,5,2,3,3,10,4,5,4,12,15,10,17,15,18,16,14,9,10,11,18,17,18,16,17,18)
```