---
title: "GEOG606 - HW1"
author: "Anika Cartas"
date: "Saturday, February 14, 2015"
output: html_document
---

####1. Solve the following linear equations:

-x + 2y = 1

3x +  y = 3

```{r}

left = matrix(c(-1,3,2,1,1,3),2,2); left
right = c(1,3); right 
solve(left,right)

```

If we add another equation, x + y = 2, is there a solution for the set of 3 equations? 

  * No. We can only solve for a square matrix.

If not, could you give a best estimate of x and y that satisfy the equations?

  * **Not sure! Get help on this one!**

####2. Calculate the covariance matrix of the following 3*2 data matrix:

```{r}
x = matrix(c(1,2,3,3,2,1),3,2); x
cov(x)
```

What does the covariance tell you about the data?

 * The covariance matrix shows the variances of the two variables (x,y) along the diagonal, and the covariances between them in the other positions.

 * In this case, the variance of both x and y is 1, while their covariance is -1, meaning that as x grows y shrinks, and vice versa.

####3.Use the following dataset to demonstrate the Central Tendency Theorem:

```{r}
data = c(3,2,3,4,3,2,5,4,5,2,3,3,10,4,5,4,12,15,10,17,15,18,16,14,9,10,11,18,17,18,16,17,18)
```

3.1. Randomly select 5 values and calculate their mean. 

```{r}
s1 = sample(data,5); s1
mean(s1)
```

Repeat 3 times, and calculate the mean of the sampling mean. Then, repeat 5, 10, 20, and 50 times. 

```{r}
takeSamples <- function(data,sizes, sampleSize){
  allMeans = matrix(NA,1,2)
  for (n in sizes){
    curMeans = vector()
    for (i in c(1:n)){
      curMeans = c(curMeans,mean(sample(data,sampleSize)))
    }
    allMeans = rbind(allMeans,c(n,mean(curMeans)))
  }
  allMeans = allMeans[-1,]
  colnames(allMeans) = c("Times.Sampled","Mean")  
  allMeans #return value
}

means.5 = takeSamples(data,c(3,5,10,20,50),sampleSize=5); means.5
```

Use a plot to show how the mean of the sampling mean changes as the number of repeats increases.

```{r}

#Out of curiosity, plotting larger sizes of samples (first, so that axes match up)
means.5.1000 = takeSamples(data,seq(1,1000,5),sampleSize=5)
plot(means.5.1000, main="Means of Various Samples of 5 from Dataset", col="grey", xlab="Times Sampled")

#Request from assignment in red
points(means.5,col="red")

mtext("With a large enough number of samples, convergence is noticeable.")

```

3.2. Repeat 3.1. but select 10 values during each random selection. Explain these plots in relation to the Central Tendency Theorem.

```{r}
means.10.1000 = takeSamples(data,seq(1,1000,10),sampleSize=5)

plot(means.10.1000, main="Means of Various Samples of 10 from Dataset", col="grey", xlab="Times Sampled")

means.10 = takeSamples(data,c(3,5,10,20,50),sampleSize=5)
points(means.10, col="red")

mtext("A similar convergence to the previous example is noticeable.")
```

3.3. How does the sampling mean approach normal distribution in these experiments? Use histograms to show.

```{r}

hist(means.5, col=rgb(0,0,1,0.5), ylim=c(0,10), main="Distribution of Means of Means", xlab="Mean")
hist(means.10, col=rgb(0,1,0,0.5), add=T)

legend("topright", c("Sample(5)","Sample(10)"),col=c(rgb(0,0,1,0.5),rgb(0,1,0,0.5)), lwd=10)
box() #decorative box around data
mtext("Difficult to note anything with this small a set of samples.")


```

