---
title: "GEOG606 - HW04"
author: "Anika Cartas"
date: "Monday, March 30, 2015"
output: html_document
---

### Assignment Description

A dataset consisting different levels of wetland inundation measurements and spectral values from Landsat and multiple spectral bands is available at ELMS. Use this dataset to:
 
```{r}
data <- read.csv("WetlandInundation_Landsat.csv")

```

###1. Evaluate the correlations 

Evaluate the correlations between wetland inundation and individual bands and spectral indices

```{r}
WI.cor <- apply(data[,2:17], 2, function(x) {
  cor(data$Wetland.Inundation, x, method="pearson")
})

WI.cor <- sort(WI.cor)

```

###2. Develop stepwise linear regression models 

Develop stepwise linear regression models between inundation and the Landsat data using both forward selection and backward elimination methods. 

```{r}
#Forward Variable Selection: F-tests
attach(data)

add1(lm(Wetland.Inundation~1), as.formula(paste("Wetland.Inundation ~", paste(names(data)[2:17],collapse="+"))), test="F")

#Pick minimum AIC as next item to add.
add1(lm(Wetland.Inundation ~ TCW.TCG), as.formula(paste("Wetland.Inundation ~", paste(names(data)[2:17],collapse="+"))), test="F")

#Only one significant to pick from, IVR
add1(lm(Wetland.Inundation ~ TCW.TCG + IVR), as.formula(paste("Wetland.Inundation ~", paste(names(data)[2:17],collapse="+"))), test="F")

#Add Band4
add1(lm(Wetland.Inundation ~ TCW.TCG + IVR + Band4), as.formula(paste("Wetland.Inundation ~", paste(names(data)[2:17],collapse="+"))), test="F")

#Add NDSI
add1(lm(Wetland.Inundation ~ TCW.TCG + IVR + Band4 + NDSI), as.formula(paste("Wetland.Inundation ~", paste(names(data)[2:17],collapse="+"))), test="F")

#Compare with automatic R version:
step(lm(Wetland.Inundation ~ 1), as.formula(paste("Wetland.Inundation ~", paste(names(data)[2:17],collapse="+"))), direction="forward")

```

Discuss your variable selection decisions.

