---
title: "Homework 05"
author: "Anika Cartas"
date: "Wednesday, April 08, 2015"
output: word_document
---

###0. Read Data

The table lists measurements of DC output at different wind velocity.

* Dependent variable: DC
* Independent variable: Wind Velocity

```{r}
data <- data.frame(
  vel = c(5,6,3.4,2.7,10,9.7,9.55,3.05,8.15,6.2,2.9,6.35,4.6,5.8,7.4,3.6,7.85,8.8,7,5.45,9.1,10.2,4.1,3.95,2.45),
  dc = c(1.582,1.822,1.057,0.5,2.236,2.386,2.294,0.558,2.166,1.866,0.653,1.93,1.562,1.737,2.088,1.137,2.179,2.112,1.8,1.501,2.303,2.310,1.194,1.144,0.123))

attach(data)
```

###1. Linear Regression

```{r}
reg.linear <- lm(dc~vel)

plot(dc~vel, main="Linear Regression")
abline(reg.linear, col="red")
```

###2. Polynomial Regression

```{r}
plot(dc~vel, main="Polynomial Regression")
lines(lowess(dc~vel), col="green")
legend("bottomright", c("Estimate","Modelled"), col=c("green","darkgreen"), lwd=2)
```

Using the lowess function, we visualize this to be a 4th degree polynomial.

```{r}
reg.poly <- lm(dc ~ poly(vel, degree=4, raw=TRUE))
lines(vel, predict(reg.poly), col="darkgreen") #Something is going very wrong here!!!
```

###3. Transformation Based Regression

The Box-Cox Power Transformation identifies an appropriate exponent (lambda) to transform the data into a normal shape. The Lambda value indicates the power to which all data should be raised. 

When lambda is plotted against log likelihood, the best lambda is the one that maximizes the log likelihood.

```{r}
library(MASS)
bc <-boxcox(dc ~ vel)

(maxll <- max(bc$y))
(lambda <- bc$x[which(bc$y == maxll)])
```

Maximum log likelihood of 17.65 is achieved when lamda is equal to 2.

```{r}
reg.trans<-lm(dc^2~vel)
plot(dc^2 ~ vel, main="Transformation Based Regression")
abline(reg.trans, col="blue")
```

###4. Nonlinear Regression

```{r}
```

###5. Discussion

Discuss which models are best for this dataset and the physical processes described by your models.

```{r}
```
